---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Libraries
```{r echo=TRUE}
library(data.table)
library(ggplot2)
```

## Loading and preprocessing the data
```{r echo=TRUE}
#Load the data
original <- fread("activity.csv")

#NA number
naNumber <- sum(is.na(original$steps))

#Omit NA records
pmData <- na.omit(original)

```

## Data characteristics
```{r echo=TRUE}
str(pmData)
dim(pmData)
summary(pmData)
head(pmData)
```

## What is total number of steps taken per day?

#Calculate the total number of steps taken per day
```{r echo=TRUE}
stepsByDay <- pmData[,sum(steps),by = date]
```

#Make a histogram of the total number of steps taken each day
```{r echo=TRUE}
qplot(V1, data=stepsByDay)
```

#Calculate and report the mean and median of the total number of steps taken per day
```{r echo=TRUE}
mean(stepsByDay$V1)
median(stepsByDay$V1)

```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE}
fiveMinIntervalAverages <- pmData[,mean(steps),by=interval]

# Make a time series plot
with(fiveMinIntervalAverages, plot(interval, V1, type="l", ylab="steps" ))

# 5-minute interval, on average across all the days in the dataset, with the maximum number of steps.

apply(fiveMinIntervalAverages, MARGIN = 2, function(x) max(x, na.rm=TRUE))


```

## Imputing missing values
## Total number of missing values in the dataset. it is calculated at the beginning, before removing NA rows.
```{r}
naNumber
```

##Devise a strategy for filling in all of the missing values in the dataset

# Find the NA positions
```{r}
naPos <- which(is.na(original$steps))
```

# Create a vector of means
```{r}
meanVec <- rep(mean(original$steps, na.rm=TRUE), times=length(naPos))
```

# Replace the NAs with the means
original[naPos, "steps"] <- meanVec

# new dataset with means
head(original)

## Are there differences in activity patterns between weekdays and weekends?
# Create a new factor variable in the dataset with two levels - "weekday" and "weekend" 
```{r}
original$weekday <- tolower(weekdays(as.Date(original$date)))
original <- cbind(original, 
                      daytype=ifelse(original$weekday == "sábado" | 
                                     original$weekday == "domingo", "weekend", 
                                     "weekday"))
```







